---
title: "Codebook Assignment"
output:
  pdf_document: default
  html_document: default
date: "November 14, 2017"
---

This assignment is meant to direct the focus of your poster project while also exposing you to the importance of a well-documented codebook. To that end you should select only the variables you are planning to include in your potential analysis, be that a linear regression model or a data visualization. 

Your final codebook document should be clearly labelled with the header titles mentioned in the following questions. For parts 2 and 3 you should write in complete sentences as if you were writing an essay. **Do not make lists.**  There should not be any code in your knitted PDF, just the output of the `codebook` function. To turn in your final assignment commit your PDF and Rmd files to Git and push them to GitHub.  

1. Create a new R Markdown file and name it after your data, followed by the word codebook (e.g.. my in-class example might be *PSID_CDI_Codebook*).

2. In the first section of your codebook, titled **Study Design**, describe the study design including the purpose of the study, the sponsor of the study, the name of the data collection organization, and the specific methodology used including the mode of data collection, the method of participant recruitment (if any), and the length of the field period. (12 points)

3. The second section, **Sampling** should clearly document all available sampling information. This includes a description of the population, the methods used to draw the sample, and any special conditions associated with the sample (i.e groups that were oversampled). (12 points)

4. Section three should be titled **Variable Index**. Here you should utilize the sample code shown in lecture (and reproduced below) to build a `data.set` version of your dataset. You will need to install and load the `memisc` package. 

Each variable in your dataset should be given a `description` and a unit of `measurement`. If there are `labels` associated with the underlying numeric values of the data those should be specified as should any `missing.value` codes including `NA`. Lastly, if your variable is a survey item/interview question you must provide the `wording` as well. 

Once you've added the information above to your `data.set` object, make a call to `codebook()` to have your variable index printed out. (26 points)   

```{r, echo = FALSE, message = FALSE}
# Setting working directory 
setwd("H:/Teaching/Sociology 321/Assignments/Project/Codebook/")

# Loading libraries
library(memisc)
library(tidyverse)

# Read in data
data <- read_csv("psid_example.csv")
```

```{r}
# Create data.set object from "data" object (tbl)
data_set <- as.data.set(data)

# Look at new data.set object
data_set

# Creating variable index for data_set
data_set <- within(data_set,{
  # Description of the variables
  description(id) <- "Unique person identification number(1968 Interview
                      Number(ER30001)) + the Person Number(ER30002))"
  description(age) <- "Preloaded age of child from coverscreen (in years)"
  description(cdi) <- "Child Depression Index"
  description(sad) <- "Sad"
  description(okay) <- "Do Things Okay"
  description(selfhate) <- "Hate Myself"
  description(cry) <- "Crying"
  description(irritability) <- "Bother Me"
  description(appearance) <- "Look Okay"
  description(isolation) <- "Feel Alone"
  description(friends) <- "Plenty of Friends"
  description(loved) <- "Loves Me"
  description(work_out) <- "Things Will Work Out"
  
  # Wording of survey item/interview questions (if applicable)
  wording(sad) <- "Select the number next to the a sentence from the group that 
  best describes your feelings during the last two weeks."
  wording(okay) <- "Select the number next to the a sentence from the group that
  best describes your feelings during the last two weeks."
  wording(selfhate) <- "Select the number next to the a sentence from the group 
  that best describes your feelings during the last two weeks."
  wording(cry) <- "Select the number next to the a sentence from the group that 
  best describes your feelings during the last two weeks."
  wording(irritability) <- "Select the number next to the a sentence from the 
  group that best describes your feelings during the last two weeks."
  wording(appearance) <- "Select the number next to the a sentence from the 
  group that best describes your feelings during the last two weeks."
  wording(isolation) <- "Select the number next to the a sentence from the 
  group that best describes your feelings during the last two weeks."
  wording(friends) <- "Select the number next to the a sentence from the group 
  that best describes your feelings during the last two weeks."
  wording(loved) <- "Select the number next to the a sentence from the group 
  that best describes your feelings during the last two weeks."
  wording(work_out) <- "Select the number next to the a sentence from the group 
  that best describes your feelings during the last two weeks."
  
  # type of measurement
  measurement(id) <- "interval"
  measurement(age) <- "ratio"
  measurement(cdi) <- "ordinal"
  measurement(sad) <- "ordinal"
  measurement(okay) <- "ordinal"
  measurement(selfhate) <- "ordinal"
  measurement(cry) <- "ordinal"
  measurement(irritability) <- "ordinal"
  measurement(appearance) <- "ordinal"
  measurement(isolation) <- "ordinal"
  measurement(friends) <- "ordinal"
  measurement(loved) <- "ordinal"
  measurement(work_out) <- "ordinal"
  
  # labels associated with underlying numeric
  labels(sad) <- c(
    "I am sad once in a while"                      =  1,
    "I am sad many times"                           =  2,
    "I am sad all the time"                         =  3,
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  labels(okay) <- c(
    "I do most things O.K."                         =  1,
    "I do many things O.K."                         =  2,
    "I do everything wrong"                         =  3,
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  labels(selfhate) <- c(
    "I hate myself"                                 =  1,
    "I do not like myself"                          =  2,
    "I like myself"                                 =  3,
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  labels(cry) <- c(
    "I feel like crying every day"                  =  1,
    "I feel like crying many days"                  =  2,
    "I feel like crying once in awhile"             =  3,
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  labels(irritability) <- c(
    "Things bother me all the time"                 =  1,
    "Things bother me many times"                   =  2,
    "Things bother me once in a while"              =  3,
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  labels(appearance) <- c(
    "I look O.K."                                   =  1,
    " There are some bad things about my looks"     =  2,
    "I look ugly          "                         =  3,
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  labels(isolation) <- c(
    "I do not feel alone"                           =  1,
    "I feel alone many times"                       =  2,
    "I feel alone all the time"                     =  3,
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  labels(friends) <- c(
    "I have plenty of friends"                      =  1,
    "I have some friends, but I wish I had more"    =  2,
    "I do not have any friends"                     =  3,
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  labels(loved) <- c(
    "Nobody really loves me"                        =  1,
    "I am not sure if anybody loves me"             =  2,
    "I am sure that somebody loves me"              =  3,
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  labels(work_out) <- c(
    "Nothing will ever work out for me"             =  1,
    "I am not sure if things will work out for me"  =  2,
    "Things will work out for me O.K."              =  3,
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  
  # any annotation/notes you wish to add
  annotation(cdi)["Note"] <- "This variable is the raw score of the CDI. Items
  `sad`, `okay`, `selfhate`, `irritability`, `friends` were recoded to: 1=0; 
  2=1; 3=2 and items `cry`, appearance`, `isolation`, `loved`, and `work_out` 
  were recoded to: 1=2; 2=1; 3=0. The 10 items were then summed if valid data 
  points were available on all 10 items."
  
  # How are missing values coded?
  missing.values(cdi) <- c(8, 9, 0)
  missing.values(sad) <- c(8, 9, 0)
  missing.values(okay) <- c(8, 9, 0)
  missing.values(selfhate) <- c(8, 9, 0)
  missing.values(cry) <- c(8, 9, 0)
  missing.values(irritability) <- c(8, 9, 0)
  missing.values(appearance) <- c(8, 9, 0)
  missing.values(isolation) <- c(8, 9, 0)
  missing.values(friends) <- c(8, 9, 0)
  missing.values(loved) <- c(8, 9, 0)
  missing.values(work_out) <- c(8, 9, 0)
  
})
```

\pagebreak

```{r}
codebook(data_set)
```

\pagebreak

# Shorter, more efficient version of code above

```{r, eval = FALSE}
# Creating variable index for data_set (truncated, more efficient code)
data_set <- within(data_set,{
  # Description of the variables
  description(id) <- "Unique person identification number (1968 Interview
                      Number(ER30001) + the Person Number(ER30002))"
  description(age) <- "Preloaded age of child from coverscreen (in years)"
  description(cdi) <- "Child Depression Index"
  description(sad) <- "Sad"
  description(okay) <- "Do Things Okay"
  description(selfhate) <- "Hate Myself"
  description(cry) <- "Crying"
  description(irritability) <- "Bother Me"
  description(appearance) <- "Look Okay"
  description(isolation) <- "Feel Alone"
  description(friends) <- "Plenty of Friends"
  description(loved) <- "Loves Me"
  description(work_out) <- "Things Will Work Out"
  
  # Wording of survey item/interview questions (if applicable)
  foreach(x=c(sad, okay, selfhate, cry, irritability, appearance, 
              isolation, friends, loved, work_out),{
    wording(x) <- "Select the number next to the a sentence from the group that 
    best describes your feelings during the last two weeks."
  })

  # type of measurement
  measurement(id) <- "interval"
  measurement(age) <- "ratio"
  foreach(x=c(cdi, sad, okay, selfhate, cry, irritability, appearance, 
              isolation, friends, loved, work_out),{
    measurement(x) <- "ordinal"
  })
  
  # labels associated with underlying numeric
  labels(sad) <- c(
    "I am sad once in a while"                      =  1,
    "I am sad many times"                           =  2,
    "I am sad all the time"                         =  3)
  labels(okay) <- c(
    "I do most things O.K."                         =  1,
    "I do many things O.K."                         =  2,
    "I do everything wrong"                         =  3)
  labels(selfhate) <- c(
    "I hate myself"                                 =  1,
    "I do not like myself"                          =  2,
    "I like myself"                                 =  3)
  labels(cry) <- c(
    "I feel like crying every day"                  =  1,
    "I feel like crying many days"                  =  2,
    "I feel like crying once in awhile"             =  3)
  labels(irritability) <- c(
    "Things bother me all the time"                 =  1,
    "Things bother me many times"                   =  2,
    "Things bother me once in a while"              =  3)
  labels(appearance) <- c(
    "I look O.K."                                   =  1,
    "There are some bad things about my looks"     =  2,
    "I look ugly          "                         =  3)
  labels(isolation) <- c(
    "I do not feel alone"                           =  1,
    "I feel alone many times"                       =  2,
    "I feel alone all the time"                     =  3)
  labels(friends) <- c(
    "I have plenty of friends"                      =  1,
    "I have some friends, but I wish I had more"    =  2,
    "I do not have any friends"                     =  3)
  labels(loved) <- c(
    "Nobody really loves me"                        =  1,
    "I am not sure if anybody loves me"             =  2,
    "I am sure that somebody loves me"              =  3)
  labels(work_out) <- c(
    "Nothing will ever work out for me"             =  1,
    "I am not sure if things will work out for me"  =  2,
    "Things will work out for me O.K."              =  3)
  foreach(x=c(sad, okay, selfhate, cry, irritability, appearance, 
              isolation, friends, loved, work_out),{
    labels(x) <- c(
    "Don't know"                                    =  8,
    "Not applicable; Answer refused"                =  9,
    "Inappropriate; child age 8-11"                 =  0)
  })
  
  # any annotation/notes you need to add to further explicate certain variables
  annotation(cdi)["Note"] <- "This variable is the raw score of the CDI. Items
  `sad`, `okay`, `selfhate`, `irritability`, `friends` were recoded to: 1=0; 2=1;
  3=2 and items `cry`, appearance`, `isolation`, `loved`, and `work_out` were
  recoded to: 1=2; 2=1; 3=0. The 10 items were then summed if valid data points 
  were available on all 10 items."
  
  # How are missing values coded?
  foreach(x=c(cdi, sad, okay, selfhate, cry, irritability, appearance, 
              isolation, friends, loved, work_out),{
    missing.values(x) <- c(8, 9, 0)
  })
  
})

```


