---
title: "Lab 10 - Merging Data"
author: "Snow Christensen"
date: "November 2, 2017"
output:
  pdf_document: default
  html_document: default
---

Using your own dataset (which may include more than one table) carry out the following data cleaning steps. Knit together the PDF document and commit both the Lab 10 RMD file and the PDF document to Git. Push the changes to GitHub so both documents are visible in your public GitHub repository. 

1. For your poster project, do you have multiple tables you'd like to join together to create your complete dataset? If so, describe what each table represents. 

No, I will just be using one. 


2. What is/are your primary key(s)? If you have more than one table in your data, what is/are your foreign key(s)? Do your primary key(s) and foreign key(s) have the same name? If not, what does this mean for the way you need to specify potential data merges?

My primary keys are RACE, NOASSER, PLEASE, AUTHSI and PRESSSI.

3. If you do not need to merge tables to create your final dataset, create a new dataset from your original dataset with a `grouped_by()` summary of your choice. You will use this separate dataset to complete the following exercises. 

```{r}
#load package to read .por files
library("haven")

#load package tidyverse for the rest of the lab
install.packages("tidyverse")
library("tidyverse")

#read in full dataset
data <- read_por("~/Projects/sexual_violence_college/ICPSR_03212/DS0001/03212-0001-Data.por")

#convert original dataset to a tbl
as_tibble(data)

#read in data_condensed from last lab
data_condensed <- data %>%
  select(RACE, NOASSERT, PLEASE, AUTHSI, PRESSSI)

# create first new data subset
practice_data1 <- head(data_condensed, n=15)

# create second new data subset
practice_data2 <- tail(data_condensed, n=15)

#group practice_data1 by SEXACTS variable and summarise the min and max of that variable
practice_data1 %>% 
  group_by(NOASSERT) %>%
    summary(min = min(SEXACTS),
            max = max(SEXACTS))



```
If you are merging separate tables as part of your data manipulation process, are your keys of the same data type? If not, what are the differences? Figure out the appropriate coercion process(es) and carry out the steps below. 

I am not merging separate tables for my real data, but for this practice they are of the same data type. 

4. Perform each version of the mutating joins (don't forget to specify the `by` argument) and print the results to the console. Describe what each join did to your datasets and what the resulting data table looks like. For those joining two separate datasets, did any of these joins result in your desired final dataset? Why or why not?

The full_join() function did because it added all of the rows from the corresponding columns in the secondary data subset to the first. It allowed me to create a more complete final dataset than the other joins did, which is what I wanted.
```{r}


#perform left join with practice_data1 as pirmary dataset
left_join(practice_data1, practice_data2, by = c("RACE", "NOASSERT", "PLEASE", "AUTHSI", "PRESSSI"))

#perform right join
right_join(practice_data1, practice_data2, by = c("RACE", "NOASSERT", "PLEASE", "AUTHSI", "PRESSSI"))

#perform inner join 
inner_join(practice_data1, practice_data2, by = c("RACE", "NOASSERT", "PLEASE", "AUTHSI", "PRESSSI"))

#perform full join
full_join(practice_data1, practice_data2, by = c("RACE", "NOASSERT", "PLEASE", "AUTHSI", "PRESSSI"))

```

5. Do the same thing with the filtering joins. What was the result? Give an example of a case in which a `semi_join()` or an `anti_join()` might be used with your primary dataset

semi_join() could be used to find similarities between the two separate datasets and anti_join() could be used to find the differences. 

```{r}

#perform semi_join()
semi_join(practice_data1, practice_data2, by = c("RACE", "NOASSERT", "PLEASE", "AUTHSI", "PRESSSI"))

#perform anti_join()
anti_join(practice_data1, practice_data2, by = c("RACE", "NOASSERT", "PLEASE", "AUTHSI", "PRESSSI"))
```
6. What happens when you apply the set operations joins to your tables? Are these functions useful for you for this project? Explain why or why not. If not, give an example in which one of them might be usefully applied to your data. 

```{r}
#use union() function to return every row in both datasets
union(practice_data1, practice_data2, by = c("RACE", "NOASSERT", "PLEASE", "AUTHSI", "PRESSSI"))
```

7. If you have any reason to compare tables, apply `setequal()` below. What were the results? 
```{r}

setequal(practice_data1, practice_data2, by = c("RACE", "NOASSERT", "PLEASE", "AUTHSI", "PRESSSI"))
```

The results tell me if the tables are equal and if they are not, how they compare. For mine they are not equal and it lists all of the rows in x but not y and vice versa.


8. What is the purpose of binding data and why might you need to take extra precaution when carrying out this specific form of data merging? If your data requires any binding, carry out the steps below and describe what was accomplished by your merge.

Binding data is useful to add one dataset to the bottom of another if the variables are exactly the same. You have to be careful though because r doesn't double check that the variables are the same and if they are off by a little it could mess up your entire dataset.


9. Do you need to merge multiple tables together using the same type of merge? If so, utilize the `reduce()` function from the `purrr` package to carry out the appropriate merge below. 

I don't need to merge multiple tables.


10. Are there any other steps you need to carry out to further clean, transform, or merge your data into one, final, tidy dataset? If so, describe what they are and carry them out below.

I don't think so, my data is already very tidy.










