---
title: "sexual_violence_codebook"
author: "Snow Christensen"
date: "November 7, 2017"
output: pdf_document
---



## Study Design ##
The study was a longitudinal study of women and men in stats-supported universities completed over the course of five years with the same respondents in two different cohorts. It is a qualitative study that had respondents complete one survey in groups at the beginning of the year and then a follow up survey at the end of the year. For this analysis I will be using the female data from the first survey about the respondents' experiences of sexual assault from 14-18 years old and what kind of person they think they are. 

## Sampling ##
This was conducted using a convenience sample of men and women of the undergraduate classes of 1990-1995. For the specific survey I am analyzing they had an 83% response rate.

## Variable Index ##


```{r, echo = FALSE, message = FALSE}
#install.packages("knitr")
library("knitr")
# Loading libraries
#install.packages("memisc", repos="https://CRAN.R-project.org/package=memisc")
library("memisc")
library("tidyverse")
library("haven")
library("dplyr")

# Read in data
data <- haven::read_por("~/Projects/sexual_violence_college/ICPSR_03212/DS0001/03212-0001-Data.por")

data <- as_tibble(data)

#create data subset I want to use
data_condensed <- data %>% 
  dplyr::select(RACE, NOASSERT, PLEASE,  AUTHSI, PRESSSI)

#convert columns to character
data_condensed$RACE <- as.character(data_condensed$RACE)

data_condensed$NOASSERT<- as.character(data_condensed$NOASSERT)

data_condensed$PLEASE <- as.character(data_condensed$PLEASE)

data_condensed$AUTHSI <- as.character(data_condensed$AUTHSI)

data_condensed$PRESSSI <- as.character(data_condensed$PRESSSI)


```

```{r}
data_set <- data.set(
                     RACE = sample(data_condensed$RACE),
                     NOASSERT = sample(data_condensed$NOASSERT),
                     PLEASE = sample(data_condensed$PLEASE),
                     AUTHSI = sample(data_condensed$AUTHSI),
                     PRESSSI = sample(data_condensed$PRESSSI)
                     )

# Look at new data.set object
data_set

# Creating variable index for data_set
data_set <- within(data_set,{
  # Description of the variables
  description(RACE) <-"What was the race of the person?"
  description(NOASSERT) <- "I stay in the background and let other people tell me what to do."
  description(PLEASE) <- "I try to please people and make them like me by giving in to them."
  description(AUTHSI) <- "Intercourse because of authority."
  description(PRESSSI) <- "Intercourse because of pressure."
  
  # Wording of survey item/interview questions
  wording(RACE) <- "What is your race or ethnic background?"
  wording(NOASSERT) <- "The questions below ask you to say what kind of a person you think you are. For example, here is a statement: 'I watch TV all the time.' If you watch TV all the time, that sentence is very much like you. If you watch TV a lot, then the sentence is mostly like you. Or you may watch TV only a little bit--then the sentence is a little like you."
  wording(PLEASE) <- "The questions below ask you to say what kind of a person you think you are. For example, here is a statement: 'I watch TV all the time.' If you watch TV all the time, that sentence is very much like you. If you watch TV a lot, then the sentence is mostly like you. Or you may watch TV only a little bit--then the sentence is a little like you."
  wording(AUTHSI) <- "Have you engaged in sexual intercourse when you didn't want to because a male used his position of authority (boss, teacher, camp counselor, supervisor) to make you?"
  wording(PRESSSI) <- "Have you given in to sexual intercourse when you didn't want to because you were overwhelmed by a male's continual arguments and pressure?"

  
  # type of measurement
  measurement(RACE) <- "nominal"
  measurement(NOASSERT) <- "nominal"
  measurement(PLEASE) <- "nominal"
  measurement(AUTHSI) <- "nominal"
  measurement(PRESSSI) <- "nominal"

  
  # labels associated with underlying numeric
  labels(RACE) <- c(
    "No response"                                   =  0,
    "White"                                         =  1,
    "Black"                                         =  2,
    "Other"                                         =  3,
    "Missing"                                       =  9)
  
  labels(NOASSERT) <- c(
    "No response"                                   =  0,
    "Not at all like me"                            =  1,
    "A little like me"                              =  2,
    "Somewhat like me"                              =  3,
    "Mostly like me"                                =  4,
    "Very much like me"                             =  5,
    "Missing"                                       =  9)
  
  labels(PLEASE) <- c(
    "No response"                                   =  0,
    "Not at all like me"                            =  1,
    "A little like me"                              =  2,
    "Somewhat like me"                              =  3,
    "Mostly like me"                                =  4,
    "Very much like me"                             =  5,
    "Missing"                                       =  9)
  
  labels(AUTHSI) <- c(
    "No response"                                   =  0,
    "Never"                                         =  1,
    "At least once"                                 =  2,
    "Missing"                                       =  9)
  
  labels(PRESSSI) <- c(
    "No response"                                   =  0,
    "Never"                                         =  1,
    "At least once"                                 =  2,
    "Missing"                                       =  9)
  
  # How are missing values coded?
  missing.values(RACE) <- c(0, 9)
  missing.values(NOASSERT) <- c(0, 9)
  missing.values(PLEASE) <- c(0, 9)
  missing.values(AUTHSI) <- c(0, 9)
  missing.values(PRESSSI) <- c(0, 9)
  
})
```

\pagebreak

```{r}

codebook(data_set)
```







